---
title: "Mini-Project 2"
author: "Misha Ritoch and Megan Pak"
date: "March 26, 2017"
output: html_document
---

```{r, include=FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```


The 2012 election was estimated to cost over $6 billion, going down in history as the most expensive election at the time. With such an expensive and competitive race among two extremely polarized political parties, our group thought it would be interesting to analyze which parties were able to raise more money and figure out how this played into the election. Using data from the Federal Election Commission, we aimed to determine which parties raised more money across all 50 states and why that might be. We used two different data sets from the Federal Election Commission--”candidates” and “contributions”--to create a set of two bar plots, first analyzing the total donation contributions across all 50 states and then analyizing the top two states with the highest donation contributions. 



```{r, warning=FALSE, message=FALSE}
library(dplyr) 
library(tidyr) 
candidates %>%
  group_by(cand_state)
contr_data <- contributions %>% inner_join(candidates, by = "cand_id")
```

```{r, warning=FALSE, message=FALSE}
party_money <- contr_data %>%
  group_by(cand_state, cand_party_affiliation) %>%
  summarize(
    N=n(), total = sum(transaction_amt)) %>%
  filter(cand_party_affiliation %in% c("REP", "DEM"))
```

```{r, warning=FALSE, message=FALSE}
party_money %>%
  arrange(desc(total))
```

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(ggthemes)
ggplot(party_money, aes(x=cand_party_affiliation, y=total)) + 
                 geom_bar(stat="identity", aes(color = cand_party_affiliation, fill=cand_party_affiliation)) +
  facet_wrap(~cand_state, ncol = 10) + 
  scale_y_continuous(labels = scales::comma) +
scale_fill_manual(values = c("DEM" = "#0000FF", "REP" = "#FF0000")) +
scale_color_manual(values = c("DEM" = "#0000FF", "REP" = "#FF0000")) +
  theme_few() +
  xlab("Party ID") + 
  ylab("Total Contributions ($)") + 
  ggtitle("Figure 1. Total Contributions for all 50 States, 2012 Election")
```


After joining the "candidates" and "contributions" data sets, we were able to create barplots with the contributions made to Democrats and Republicans for each state. From Figure 1, it is clear that the states with the highest number of donation contributions include Massachusetts (MA) and Illinois (IL), home state to the 2012 presidential nominees, Mitt Romney and Barack Obama. After Massachusetts and Illinois, California, Texas and New York ranked among the next largest contributors to the presidential campaign. This is probably indicative of top donors living in large cities. California and New York are predominantly Democratic with Texas generally identifying as Republican, which is also evident in Figure 1. Additionally, it is interesting to see that typical swing states were among the states with some of the largest donation contributions. These states included Florida, Virginia, Ohio, Pennsylvania, and Wisconsin. 



```{r, warning=FALSE, message=FALSE}
MAIL <- party_money %>% 
  filter(cand_state %in% c("MA", "IL"))
```

```{r, warning=FALSE, message=FALSE}
ggplot(MAIL, aes(x=cand_party_affiliation, y=total)) + 
                 geom_bar(stat="identity", aes(color = cand_party_affiliation, fill=cand_party_affiliation)) +
  facet_wrap(~cand_state, ncol = 2) + 
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual(values = c("DEM" = "#0000FF", "REP" = "#FF0000")) +
  scale_color_manual(values = c("DEM" = "#0000FF", "REP" = "#FF0000")) +
  theme_few() +
  xlab("Party ID") + 
  ylab("Total Contributions ($)") + 
  ggtitle("Figure 2. Massachusetts and Illinois Contribution, ELection 2012")
```


Figure 2 looks at Massachusetts and Illinois side by side, which had the highest donation contributions. In Illinois, the total contribution towards the Democratic Party was roughly $429,046,963. Conversely, Massachusetts contributed roughly $199,441,933 towards the Republican Party. In the end, Obama had an estimated $1 billion raised towards his campaign with Romney following close behind. Obama was ultimately better funded because he was able to recieve small donations, collecting donations from millions of supporters who had given up to a few dollars each. Romney, by contrast, specifically relied on large donors, including wealthy donors who gave the maximum-allowed donation and establishments who were eager to defeat Obama. In the end, Obama was able to raise and spend more money than his Republican opponent, and ultimately won the White House as well. 
